---
title: "Visualize Data"
author: "Alexis Zarate"
date: today
format: gfm
execute: 
  echo: true
  eval: true
---


<!-- If you get the error "Error in list2(na.rm = na.rm, ...) : object 'ffi_list2' not found" do the following in the Console run:  

    remove.packages("rlang")

Then Restart R to be safe. Then run:

    install.packages("rlang")

and again restart R to be safe. That should fix the problem (which is a package dependency issue).   -->



Try your code again
```{r}
#| label: setup
#| include: false

library(tidyverse)

```



## Your Turn 0

Add a setup chunk that loads the tidyverse packages.


```{r}
#| label: mpg-data
### viewing the first 6 rows of data from the mpg dataset, using head()
### dataset is 234 x 11
dim(mpg) 
head(mpg)
```


## Your Turn 1

Run the code on the slide to make a graph. Pay strict attention to spelling, capitalization, and parentheses!

```{r}
#| label: scatter-plot
### initializing a plot with ggplot(), adding layers with geom functions 
ggplot(mpg) +
  geom_point(aes(x = displ, 
                 y = hwy))
```


## Your Turn 2

Replace this scatterplot with one that draws boxplots. Use the cheatsheet. Try your best guess.

```{r}
#| label: boxplot
### utilizing a boxplot to visualize the distribution of the mpg dataset
ggplot(mpg) +
  geom_boxplot(aes(x = class, 
                   y = hwy))
```


## Your Turn 3

Make a histogram of the `hwy` variable from `mpg`. Hint: do not supply a y variable.

```{r}
#| label: histogram
### histograms used for visualizing the frequency distribution of the mpg dataset
ggplot(mpg) +
  geom_histogram(aes(x = hwy))
```

## Your Turn 4

Use the help page for `geom_histogram` to make the bins 2 units wide.

```{r}
#| label: histogram-2
### histograms used for visualizing the frequency distribution of the mpg dataset
ggplot(mpg) +
  geom_histogram(aes(x = hwy), 
                 binwidth = 2)
```


## Your Turn 5

Add `color`, `size`, `alpha`, and `shape` aesthetics to your graph. Experiment.  

```{r}
#| label: Color-to-histogram
### using 
ggplot(mpg) +
  geom_histogram(aes(x = hwy),
                    binwidth = 2,
                    alpha = 0.5,
                    fill = "#00e6e6",
                    color = "black")
```


```{r}
#| label: scatterplot-2
### scatterplot with aesthetics 
ggplot(mpg) +
  geom_point(aes(x = displ,
                 y = hwy,
                 color = class)) + 
  labs(x = "Engine Size Displacement",
       y = "Highway Mileage (mpg)",
       color = "Class of Vehicle",
       title = "The bigger the engine the less mileage we get on the highway") + 
  theme_light()
```




## Help Me

What do `facet_grid()` and `facet_wrap()` do? 
(run the code, interpret, convince your group)

```{r}
#| label: facet-functions
# Makes a plot that the commands below will modify
q <- ggplot(mpg) + geom_point(aes(x = displ, y = hwy))

q + facet_grid(. ~ cyl) ### make one row, four columns by cyl
q + facet_grid(drv ~ .) ### makes three rows, one column by drv
q + facet_grid(drv ~ cyl) ### makes three rows by drv, four columns by cyl
q + facet_wrap(~ class) ### makes a graph for each value in class
```


## Your Turn 6

Make a bar chart `class` colored by `class`. Use the help page for `geom_bar` to choose a "color" aesthetic for class.

```{r}
#| label: barplot
### barplot with aesthetics 
ggplot(mpg) +
  geom_bar(aes(x = class,
               fill = drv),
  position = position_dodge())+
  guides(fill = "none") +
  labs(x = "Class of Vehicle",
       y = "Number of Vehicles in Sample") + 
  theme_bw()
```

## Quiz

What will this code do?

```{r}
#| label: Quiz
ggplot(mpg, aes(x = displ, y = hwy)) + 
  geom_point(aes(color = class)) +
  geom_smooth(color = "black",
                  se = FALSE,
                  method = "lm")


#ggsave("example.jpg", width = 6, height = 4)
```


***

# Take aways

You can use this starter code template to make thousands of graphs with **ggplot2**.

```{r}
#| label: template
#| eval: false
#ggplot(data = <DATA>) +
  #<GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))
```